\section{Methods}
\label{sec:methods}

We describe the mathematical formulation and GPU implementation strategy for each of the five computational kernels.
Throughout, we adopt the following notation conventions: $\gamma$ denotes the Lorentzian half-width at half-maximum (HWHM) in \si{\nano\metre}, $\sigma$ denotes the Gaussian standard deviation in \si{\nano\metre}, $T$ denotes the electron temperature in \si{\electronvolt} unless subscripted as $T_K$ for Kelvin, $n_e$ denotes the electron density in \si{\per\cubic\centi\metre}, and $\lambda$ denotes wavelength in \si{\nano\metre}.
Energies $E_k$ are measured in \si{\electronvolt} above the ground state of each ionization stage, and $k_B = \SI{8.617333e-5}{\electronvolt\per\kelvin}$.

\subsection{GPU-Optimized Voigt Profile}
\label{sec:voigt}

The Voigt profile $V(\lambda)$ describes the spectral line shape arising from the convolution of Gaussian (Doppler) and Lorentzian (Stark broadening plus natural linewidth) contributions.
In terms of the Faddeeva function $w(z) = e^{-z^2} \operatorname{erfc}(-iz)$, the normalized Voigt profile is
\begin{equation}
  V(\lambda) = \frac{\operatorname{Re}[w(z)]}{\sigma \sqrt{2\pi}}, \quad
  z = \frac{(\lambda - \lambda_0) + i\gamma}{\sigma\sqrt{2}},
  \label{eq:voigt}
\end{equation}
where $\lambda_0$ is the line center wavelength.
The profile integrates to unity over wavelength and has dimensions of \si{\per\nano\metre}.

Efficient evaluation of $w(z)$ on GPU hardware requires a branch-free algorithm: conditional branches within vectorized operations cause warp divergence on NVIDIA architectures and are incompatible with JAX's automatic differentiation, which evaluates all branches of \texttt{jnp.where} during the forward pass.
The widely used \citet{Humlicek1982} W4~algorithm employs four piecewise regions selected by conditionals on $|x|$ and $y = \operatorname{Im}(z)$, making it unsuitable for JAX compilation.

We adopt the rational approximation of \citet{Weideman1994} with $N = 36$ terms.
This algorithm applies a M\"obius transformation $Z = (L + iz)/(L - iz)$ with optimal shift $L \approx 5.0454$, then evaluates a degree-35 polynomial in $Z$:
\begin{equation}
  w(z) \approx \frac{1}{\sqrt{\pi}} \frac{L}{L - iz}
    + \frac{2}{(L - iz)^2} \sum_{n=0}^{35} c_n Z^n,
  \label{eq:weideman}
\end{equation}
where the coefficients $\{c_n\}$ are precomputed constants.
This formulation is entirely branch-free, requires 236 floating-point operations per evaluation, and achieves a maximum relative error below $10^{-13}$ in 64-bit arithmetic across the entire upper half-plane $\operatorname{Im}(z) \geq 0$ \citep{Weideman1994}.
\citet{Zaghloul2024} recently proposed an improved algorithm that achieves $10^{-16}$ accuracy, which we use as an accuracy reference; however, its additional computational cost is unnecessary for LIBS applications where the Voigt profile is convolved with an instrument function of finite resolving power.
The Voigt full-width at half-maximum can be estimated from the Gaussian and Lorentzian widths via the approximation of \citet{Olivero1977}, which is accurate to 0.02\%.

For a spectrum containing $N_\mathrm{lines}$ emission lines evaluated on a wavelength grid of $N_\mathrm{wl}$ points, the full profile matrix has shape $(N_\mathrm{wl}, N_\mathrm{lines})$.
Rather than looping over lines, we broadcast the wavelength grid against the line parameter vectors:
\begin{equation}
  \mathrm{diff}_{ij} = \lambda_i - \lambda_{0,j}, \quad
  z_{ij} = \frac{\mathrm{diff}_{ij} + i\gamma_j}{\sigma_j \sqrt{2}},
  \label{eq:voigt-broadcast}
\end{equation}
where $i = 1, \ldots, N_\mathrm{wl}$ and $j = 1, \ldots, N_\mathrm{lines}$.
The entire $(N_\mathrm{wl}, N_\mathrm{lines})$ grid of Faddeeva evaluations is then executed as a single fused XLA kernel, exploiting the massive parallelism of the GPU.

\subsection{Vectorized Boltzmann Fitting}
\label{sec:boltzmann}

The Boltzmann plot method \citep{Tognoni2010} estimates the plasma temperature from emission line intensities.
For a given element~$s$ in ionization stage~$z$, plotting
\begin{equation}
  y_i = \ln\!\left(\frac{I_i \,\lambda_i}{g_{k,i}\, A_{ki,i}}\right)
  \quad \text{vs.} \quad
  x_i = E_{k,i}
  \label{eq:boltzmann-plot}
\end{equation}
yields a linear relationship $y = a + bx$ with slope $b = -1/(k_B T)$ and intercept $a = \ln(hc\, C_s\, n_e\, F_s / (4\pi\, U_z(T)))$, where $I_i$ is the measured intensity of line~$i$, $g_{k,i}$ is the upper-level statistical weight, $A_{ki,i}$ is the Einstein spontaneous emission coefficient, $E_{k,i}$ is the upper-level energy, and $F_s$ encodes the Saha ionization-stage ratios for species~$s$.
The index $i$ runs over the $N_s$ emission lines available for a given element--stage group, where $N_s$ varies from element to element (e.g., 3 lines for Li versus 40+ for Fe in a typical LIBS wavelength window).

In a conventional CF-LIBS analysis, the Boltzmann fit is performed once per element per iteration of the Saha--Boltzmann loop, typically via a general-purpose least-squares routine such as \texttt{numpy.polyfit}.
For GPU-accelerated batch computation---where thousands of plasma parameter sets must be evaluated simultaneously during manifold generation or Monte Carlo uncertainty propagation---this sequential, per-element approach becomes the bottleneck.
We describe here a formulation that fits all $D$ elements simultaneously in a single GPU kernel call, without Python loops or dynamic control flow.

\subsubsection{Closed-Form Weighted Least Squares via Scalar Accumulators}
\label{sec:boltzmann-wls}

We assign each data point a weight $w_i > 0$ (typically $w_i = 1/\sigma_{y,i}^2$, where $\sigma_{y,i}$ is the propagated intensity uncertainty) and minimize the weighted sum of squared residuals:
\begin{equation}
  \chi^2 = \sum_{i=1}^{N} w_i \bigl(y_i - a - b\, x_i\bigr)^2.
  \label{eq:boltzmann-chi2}
\end{equation}
Setting $\partial \chi^2 / \partial a = 0$ and $\partial \chi^2 / \partial b = 0$ produces the $2 \times 2$ \emph{normal equations} \citep[ch.~6]{Bevington2003}:
\begin{equation}
  \underbrace{
  \begin{pmatrix}
    S_w    & S_{wx} \\
    S_{wx} & S_{wxx}
  \end{pmatrix}}_{\displaystyle \bm{X}^\top \bm{W} \bm{X}}
  \begin{pmatrix} a \\ b \end{pmatrix}
  =
  \underbrace{
  \begin{pmatrix} S_{wy} \\ S_{wxy} \end{pmatrix}}_{\displaystyle \bm{X}^\top \bm{W} \bm{y}},
  \label{eq:normal-equations}
\end{equation}
where we have introduced five \emph{scalar accumulators}:
\begin{align}
  S_w    &= \textstyle\sum_{i=1}^{N} w_i, \quad
  S_{wx}  = \textstyle\sum_{i=1}^{N} w_i\, x_i, \quad
  S_{wy}  = \textstyle\sum_{i=1}^{N} w_i\, y_i, \notag\\
  S_{wxx} &= \textstyle\sum_{i=1}^{N} w_i\, x_i^2, \quad
  S_{wxy}  = \textstyle\sum_{i=1}^{N} w_i\, x_i\, y_i.
  \label{eq:boltzmann-sums}
\end{align}
Here $\bm{X}$ is the $N \times 2$ design matrix with rows $(1,\; x_i)$, $\bm{W} = \operatorname{diag}(w_1,\ldots,w_N)$, and $\bm{y} = (y_1,\ldots,y_N)^\top$.
For the general $p$-predictor case, solving the normal equations requires $O(p^2 N + p^3)$ operations, and numerical stability demands QR or SVD factorization of the design matrix when $p$ is large or the Gram matrix is ill-conditioned \citep[ch.~15]{Press2007}.
For $p = 2$ (intercept and slope), the system admits an exact analytic inverse via the adjugate formula, yielding:
\begin{equation}
  \begin{pmatrix} a \\ b \end{pmatrix}
  = \frac{1}{\Delta}
  \begin{pmatrix}
    S_{wxx}\, S_{wy} - S_{wx}\, S_{wxy} \\
    S_w\, S_{wxy} - S_{wx}\, S_{wy}
  \end{pmatrix},
  \label{eq:boltzmann-slope}
\end{equation}
where
\begin{equation}
  \Delta = S_w\, S_{wxx} - S_{wx}^2
  \label{eq:boltzmann-det}
\end{equation}
is the determinant of the Gram matrix.
This formulation reduces the entire least-squares solution to five inner products (the accumulators in \cref{eq:boltzmann-sums}), one determinant (3~operations), and two dot-product-style numerator evaluations (4~operations each)---a total of $10N + 11$ floating-point operations per element.
No matrix is ever explicitly constructed, inverted, or factorized.

The temperature follows from the slope as
\begin{equation}
  T = -\frac{1}{b\, k_B},
  \label{eq:boltzmann-temperature}
\end{equation}
where $k_B = \SI{8.617\,333e-5}{\electronvolt\per\kelvin}$ is the Boltzmann constant.

\subsubsection{Uncertainty Propagation from the Gram Covariance Matrix}
\label{sec:boltzmann-uncertainty}

The parameter covariance matrix is the inverse of the Gram matrix \citep[ch.~6]{Bevington2003}:
\begin{equation}
  \operatorname{Cov}(a, b) = (\bm{X}^\top \bm{W} \bm{X})^{-1}
  = \frac{1}{\Delta}
  \begin{pmatrix}
    S_{wxx}  & -S_{wx} \\
    -S_{wx}  & S_w
  \end{pmatrix},
  \label{eq:boltzmann-covariance}
\end{equation}
giving $\operatorname{Var}(b) = S_w / \Delta$ and $\operatorname{Var}(a) = S_{wxx} / \Delta$.
Because $T = -1/(b\, k_B)$, standard error propagation yields:
\begin{equation}
  \sigma_T = \left|\frac{\partial T}{\partial b}\right| \sigma_b
  = T^2\, k_B \sqrt{\frac{S_w}{\Delta}}\,,
  \label{eq:boltzmann-sigma-T}
\end{equation}
where we have used $|\partial T / \partial b| = 1/(b^2 k_B) = T^2 k_B$.
Because the five accumulators and the determinant $\Delta$ are already computed for the slope, the uncertainty requires only two additional operations (a division and a square root) per element.

The weighted coefficient of determination,
\begin{equation}
  R^2 = 1 - \frac{\sum_i w_i (y_i - a - b\, x_i)^2}{\sum_i w_i (y_i - \bar{y}_w)^2},
  \quad \bar{y}_w = S_{wy} / S_w,
  \label{eq:boltzmann-R2}
\end{equation}
quantifies the goodness of fit and serves as a diagnostic for non-LTE deviations, self-absorption, or misidentified lines.

\subsubsection{Why the Closed-Form Solution Is Preferred}
\label{sec:boltzmann-why-closed-form}

For general multi-parameter regression ($p \gg 2$), solving the normal equations directly is inadvisable because the Gram matrix condition number $\kappa(\bm{X}^\top \bm{W} \bm{X}) \approx \kappa(\bm{X})^2$ squares the condition number of the design matrix, potentially losing $2\log_{10}\kappa$ digits of precision \citep[sec.~15.4]{Press2007}.
QR or SVD factorization avoids this squaring and is the standard recommendation.
However, for $p = 2$ the condition number is bounded by the ratio of the largest to smallest weighted energy spread: $\kappa \sim S_{wxx} / S_w \sim (\Delta E_k)^2$ for typical Boltzmann data spanning 0--\SI{10}{\electronvolt}.
This gives $\kappa \sim 10^2$--$10^3$, losing 2--3 digits relative to the data---leaving 12--13 correct digits in \texttt{float64}, far exceeding the ${\sim}1$--$10\%$ physical uncertainties in LIBS line intensities.
QR or SVD factorization of a $2 \times 2$ system would add ${\sim}3\times$ computational overhead with no practical accuracy benefit.

Equally important for GPU execution: the closed-form solution is \emph{branch-free}.
Every element follows the same arithmetic path regardless of its data values, eliminating the warp divergence that would arise from iterative solvers or QR pivoting on a GPU.
The only conditional logic is a guard against degenerate fits ($|\Delta| < 10^{-30}$, indicating that all upper-level energies are identical or only one line is available), which is handled via a branchless \texttt{where} operation that returns $T = 0$ for such cases.

\subsubsection{Pad-and-Mask Batching for Variable-Length Line Lists}
\label{sec:boltzmann-pad-mask}

A multi-element CF-LIBS analysis requires fitting Boltzmann plots for $D$ element--stage groups simultaneously, where each group has a different number of usable emission lines $N_s$ (e.g., $N_{\mathrm{Li}} = 3$, $N_{\mathrm{Fe}} = 47$).
On CPU, this is handled by a Python loop over elements.
On GPU, this variable-length structure poses a vectorization challenge: standard array operations require uniform shapes along the reduction axis.

We resolve this via \emph{pad-and-mask batching}, a standard technique in GPU computing for handling ragged data:
\begin{enumerate}
  \item \textbf{Pad to uniform length.} All $D$ line lists are padded with zeros to the maximum line count $N_\mathrm{max} = \max_s N_s$, producing arrays $\bm{x}, \bm{y}, \bm{w} \in \mathbb{R}^{D \times N_\mathrm{max}}$.
  Padded positions are filled with arbitrary values (we use zero).

  \item \textbf{Construct a binary mask.}
  A mask array $\bm{m} \in \{0, 1\}^{D \times N_\mathrm{max}}$ indicates valid entries: $m_{s,i} = 1$ if line~$i$ is a real emission line for element~$s$, and $m_{s,i} = 0$ for padded positions.

  \item \textbf{Mask the weights.}
  The effective weight array is formed by element-wise multiplication:
  \begin{equation}
    \tilde{w}_{s,i} = w_{s,i} \cdot m_{s,i}.
    \label{eq:masked-weights}
  \end{equation}
  This sets $\tilde{w}_{s,i} = 0$ for all padded entries, causing them to contribute exactly zero to every accumulator sum---regardless of the $x$ and $y$ values in those positions.

  \item \textbf{Compute accumulators as matrix--vector products.}
  The five accumulators for all $D$ elements simultaneously are:
  \begin{equation}
    S_w^{(s)} = \sum_{i=1}^{N_\mathrm{max}} \tilde{w}_{s,i}, \quad
    S_{wx}^{(s)} = \sum_{i=1}^{N_\mathrm{max}} \tilde{w}_{s,i}\, x_{s,i}, \quad \ldots
    \label{eq:batched-accumulators}
  \end{equation}
  Each accumulator is a \emph{row-wise reduction} of the $(D, N_\mathrm{max})$ array, producing a vector of length~$D$.
  In JAX, this is \texttt{jnp.sum(w\_masked * x, axis=1)}.
\end{enumerate}

\noindent
The key insight is that weight-zeroing makes padding \emph{mathematically invisible}: a padded entry with $\tilde{w} = 0$ contributes $0 \cdot x = 0$ and $0 \cdot y = 0$ to every accumulator, so the slope, intercept, temperature, uncertainty, and $R^2$ are identical to those obtained from only the valid lines.
No conditional branching is needed---the mask operates entirely through multiplication.

The total operation count is $10 N_\mathrm{max} + 11$ floating-point operations per element, or $D(10 N_\mathrm{max} + 11)$ for the full batch.
For a typical multi-element LIBS spectrum with $D = 10$ elements and $N_\mathrm{max} = 50$ lines, the entire batch fit requires 5{,}110 FLOPs---well within the throughput of a single GPU warp.
The ``wasted'' computation on padded entries ($10(N_\mathrm{max} - N_s)$ operations per element) is negligible on GPU hardware, where the alternative---ragged arrays requiring per-element kernel launches or Python-level control flow---would incur orders-of-magnitude greater overhead from kernel launch latency and host--device synchronization.

\subsubsection{GPU Execution Model}
\label{sec:boltzmann-gpu}

The complete kernel is implemented as a single JAX \texttt{@jit}-compiled function.
All inputs ($\bm{x}$, $\bm{y}$, $\bm{w}$, $\bm{m}$) are \texttt{float64} arrays of shape $(D, N_\mathrm{max})$, and all outputs (slope, intercept, $T$, $\sigma_T$, $\sigma_b$, $\sigma_a$, $R^2$, $N_\mathrm{valid}$) are vectors of shape $(D,)$.
The function contains no Python loops, no conditional branches (only branchless \texttt{jnp.where} guards), and no dynamic shapes---properties required for XLA compilation to a single GPU kernel.

Because the five accumulator sums are independent reductions along axis~1 of identically shaped arrays, they execute as five parallel \texttt{sum} operations on the GPU, with the element dimension~$D$ serving as the natural batch axis.
This is not a \texttt{vmap} transformation over elements---it is a direct vectorization at the array level, where each element's fit is computed by the same arithmetic applied to different rows of the input arrays.
The branchless design ensures that all GPU threads within a warp execute the same instruction sequence, avoiding the performance penalty of warp divergence.

Enforcing \texttt{float64} precision is essential: the Gram matrix determinant $\Delta = S_w S_{wxx} - S_{wx}^2$ involves a difference of products that can suffer catastrophic cancellation in \texttt{float32} when the upper-level energies cluster near a common value.
In \texttt{float64}, the 2--3 digits lost to the condition number of the $2 \times 2$ system leave 12--13 correct digits, far exceeding what the physical measurement uncertainties warrant.

\subsection{Anderson-Accelerated Saha--Boltzmann Solver}
\label{sec:anderson}

The CF-LIBS inversion requires solving a charge-balance equation self-consistently: given a temperature $T$ and trial electron density $n_e$, the Saha equation determines ionization fractions, from which a new $n_e$ follows via charge neutrality $n_e = \sum_s \sum_{z \geq 1} z \cdot n_{s,z}$.
This defines a fixed-point problem $n_e = g(n_e)$ where $g$ encapsulates the Saha--Boltzmann system.

Simple Picard iteration $n_e^{(k+1)} = g(n_e^{(k)})$ converges at a rate governed by the spectral radius of $g'(n_e^*)$, which is typically $0.3$--$0.7$ for LIBS-relevant conditions.
Anderson acceleration \citep{Walker2011} improves upon Picard iteration by maintaining a history of the $m$ most recent iterates and their residuals $r_k = g(n_e^{(k)}) - n_e^{(k)}$, then computing an optimal linear combination:
\begin{equation}
  n_e^{(k+1)} = g(n_e^{(k)}) - (\Delta G_k + \Delta R_k)\, \bm{\gamma}_k^*,
  \label{eq:anderson}
\end{equation}
where $\Delta G_k$ and $\Delta R_k$ are matrices of iterate and residual differences over the history window, and $\bm{\gamma}_k^*$ minimizes the residual norm:
\begin{equation}
  \bm{\gamma}_k^* = \arg\min_{\bm{\gamma}} \| r_k - \Delta R_k \bm{\gamma} \|^2 + \lambda \|\bm{\gamma}\|^2.
  \label{eq:anderson-ls}
\end{equation}
The Tikhonov regularization parameter $\lambda$ stabilizes the least-squares solve when the history vectors become nearly collinear.

\citet{Toth2015} proved that Anderson acceleration with depth $m$ converges r-linearly at a rate improved by a factor depending on the gain over Picard's contraction constant.
\citet{Evans2018} showed that for sufficiently smooth fixed-point maps, depth~$m$ Anderson acceleration achieves convergence comparable to GMRES($m$) applied to the linearized residual equation.

In our JAX implementation, the iteration history is maintained as a fixed-size circular buffer within a \texttt{lax.scan} loop, ensuring a static computation graph suitable for JIT compilation.
The depth $m$ is a compile-time constant (we find $m = 1$ optimal at tolerance $10^{-6}$ for typical LIBS conditions, balancing the overhead of the least-squares solve against the iteration reduction).
When $m = 0$, the algorithm reduces exactly to Picard iteration, providing a graceful fallback.

\subsection{Compositional Closure via Exponential Simplex Mapping}
\label{sec:softmax}

CF-LIBS determines elemental compositions subject to the simplex constraint $\sum_{i=1}^D C_i = 1$ with $C_i > 0$, a structure known in statistics as compositional data \citep{Egozcue2003}.
Traditional approaches enforce this constraint via explicit normalization after each iteration, which complicates gradient-based optimization and can introduce numerical instability when compositions approach the simplex boundary.

The exponential simplex parameterization maps unconstrained parameters $\bm{\theta} \in \mathbb{R}^D$ to the probability simplex via
\begin{equation}
  C_i = \frac{\exp(\theta_i)}{\sum_{j=1}^D \exp(\theta_j)},
  \label{eq:softmax}
\end{equation}
implemented with the log-sum-exp trick for numerical stability: $\theta_i \leftarrow \theta_i - \max_j \theta_j$ before exponentiation.
The constraint $\sum_i C_i = 1$ is satisfied exactly by construction, and strict positivity $C_i > 0$ holds for all finite $\bm{\theta}$.

The Jacobian of the softmax map,
\begin{equation}
  J_{ij} = \frac{\partial C_i}{\partial \theta_j} = C_i(\delta_{ij} - C_j),
  \label{eq:softmax-jacobian}
\end{equation}
is rank $D - 1$ (the kernel is $\bm{1}$, reflecting the shift invariance $C(\bm{\theta}) = C(\bm{\theta} + c\bm{1})$) and has condition number $\kappa = 1$ at the uniform composition $C_i = 1/D$.
This Jacobian propagates gradients from any downstream loss function (e.g., spectral residual) back through the compositional constraint, enabling joint optimization of $T$, $n_e$, and $\bm{C}$ via L-BFGS-B or similar quasi-Newton methods.

\subsection{Batch Forward Model}
\label{sec:batch}

The forward model maps plasma parameters $(T, n_e, \bm{C})$ and a wavelength grid $\bm{\lambda}$ to a synthetic emission spectrum $S(\lambda)$ through four stages:
\begin{enumerate}
  \item \textbf{Saha ionization balance}: For each element~$s$, compute the ionization-stage fractions $f_z^{(s)}(T, n_e)$ from the Saha equation using partition functions $U_z(T)$ and ionization potential depression $\Delta\chi_\mathrm{DH}$.
  \item \textbf{Boltzmann level populations}: For each transition line~$l$ belonging to element~$s$ in stage~$z$, compute the upper-level number density $n_k^{(l)} = C_s \cdot n_e \cdot f_z^{(s)} \cdot g_k^{(l)} \exp(-E_k^{(l)} / T) / U_z^{(s)}(T)$.
  \item \textbf{Line emissivity}: Compute $\varepsilon_l = (hc / 4\pi\lambda_l) A_{ki}^{(l)} n_k^{(l)}$ in \si{\watt\per\cubic\metre\per\steradian} (with the \si{\per\cubic\centi\metre} to \si{\per\cubic\metre} conversion factor $10^6$).
  \item \textbf{Voigt broadening and spectrum assembly}: Evaluate the Voigt profile for each line on the wavelength grid via \cref{eq:voigt-broadcast}, then sum:
    \begin{equation}
      S(\lambda) = \sum_{l=1}^{N_\mathrm{lines}} \varepsilon_l \cdot V_l(\lambda) \quad [\si{\watt\per\cubic\metre\per\steradian\per\nano\metre}].
      \label{eq:spectrum}
    \end{equation}
\end{enumerate}

For batch computation over $B$ parameter sets $\{(T_b, n_{e,b}, \bm{C}_b)\}_{b=1}^B$---as required for manifold generation or Monte Carlo uncertainty propagation---we apply JAX's \texttt{vmap} transformation at a single level over the batch dimension:
\begin{multline}
  \texttt{batch\_spectrum} = \texttt{jit}\bigl(\texttt{vmap}\bigl(\texttt{single\_spectrum},\\
  \texttt{in\_axes}=(0, 0, 0, \text{None}, \text{None})\bigr)\bigr),
  \label{eq:vmap}
\end{multline}
where the first three axes (temperature, electron density, composition) are mapped over the batch, while the wavelength grid and atomic data are broadcast (shared) across all batch elements.

The dominant memory consumer is the profile matrix of shape
$(B,\allowbreak N_\mathrm{wl},\allowbreak N_\mathrm{lines})$.
The maximum batch size for a given GPU memory budget $M_\mathrm{GPU}$~is
\begin{equation}
  B_\mathrm{max} = \left\lfloor \frac{M_\mathrm{GPU} - M_\mathrm{overhead}}{N_\mathrm{wl} \times N_\mathrm{lines} \times d_\mathrm{bytes}} \right\rfloor,
  \label{eq:bmax}
\end{equation}
where $d_\mathrm{bytes} = 8$ for \texttt{float64} and $M_\mathrm{overhead} \approx \SI{2}{\giga\byte}$ accounts for the JAX runtime and XLA workspace.

The batch forward model achieves bit-identical results to sequential single-spectrum evaluation because \texttt{vmap} is a purely semantic transformation---it does not alter the floating-point operations within each spectrum computation, only their scheduling across GPU threads.
This property provides a strong correctness guarantee: any result obtained in batch mode is numerically identical to the same computation performed sequentially on CPU.
As reported by \citet{Kawahara2022} for ExoJAX, this \texttt{vmap}-based approach to spectral batch computation is well-suited to GPU architectures where the per-spectrum workload (dominated by the $(N_\mathrm{wl},\allowbreak N_\mathrm{lines})$ profile matrix) provides sufficient arithmetic intensity to saturate the GPU's compute units.

\subsection{Element Identification Benchmark}
\label{sec:element-id}

Element identification---determining which elements are present in the plasma before quantitative analysis can proceed---is a critical prerequisite for CF-LIBS that becomes increasingly difficult at low resolving power.
At the resolving powers typical of compact and field-deployed spectrometers (RP~$= 300$--$1000$), severe spectral overlap and line blending render peak-matching unreliable.
We present a systematic benchmark of five identification pathways on a challenging dataset of real LIBS spectra.

\subsubsection{Benchmark Dataset}

The benchmark comprises 74~LIBS spectra from the Aalto University mineral spectral library \citep{Drozdovskiy2020}: 13~pure element spectra (Al, Co, Cr, Cu, Fe, Mg, Mn, Ni, Pb, Sn, Ti, V, Zn) and 61~mineral spectra covering 45~mineral species.
% Source: element_id_benchmark_report.md §2.1
The ground truth for pure elements is the single target element; for minerals, it is the set of metallic/semi-metallic major elements from the stoichiometric formula, restricted to a 22-element search list (Fe, Ca, Mg, Si, Al, Ti, Na, K, Mn, Cr, Ni, Cu, Co, V, Li, Sr, Ba, Zn, Pb, Mo, Zr, Sn).
Light elements (O, H, C, N, S, F, Cl, B) are excluded as they produce weak or ambiguous LIBS emission in air.
The effective resolving power, estimated from isolated peak FWHM measurements, spans RP~$= 300$--$1100$ (median~$\approx 600$) over the wavelength range 200--900~nm.

\subsubsection{Scoring Methodology}

Each algorithm produces a set of detected elements per spectrum.
We compute micro-averaged precision~$P$ (identification specificity), recall~$R$ (detection sensitivity), and F$_1$ score across all 74~spectra: confirmed detections (TP, element correctly identified), spurious detections (FP, element identified but absent from the sample), and missed detections (FN, element present but not identified).
An ``exact match'' requires the detected element set to equal the ground-truth set exactly.
The spurious detection rate (FPR) is the fraction of absent elements that are incorrectly identified.
For CF-LIBS, identification specificity is more consequential than detection sensitivity: a spuriously identified element introduces extraneous lines into the Boltzmann plot, biasing the temperature determination via the slope $b = -1/(k_B T)$ and cascading systematic errors into the concentrations of all elements through the closure equation $\sum C_i = 1$.
A missed element, while it reduces the completeness of the analysis, does not corrupt the Boltzmann plots of correctly identified elements.
This asymmetry motivates our emphasis on the $F_{\beta=0.5}$ score, which weights identification specificity twice as heavily as detection sensitivity.

\subsubsection{Basis Library}

Pathways requiring basis spectra (NNLS, hybrid, forward model) share a precomputed basis library generated from Saha--Boltzmann physics \citep{Ciucci1999,Tognoni2010}.
The library covers 76~elements on a temperature grid of 4000--12{,}000~K (30~linear steps) and electron density grid of $10^{15}$--$5 \times 10^{17}$~\si{\per\cubic\centi\metre} (10~logarithmic steps), yielding 300~grid points per element.
This grid spans the plasma conditions typical of laser ablation in air: temperatures from 4000~K (late-gate, recombining plasmas) to 12{,}000~K (early-gate, highly excited plasmas), and electron densities from $10^{15}$~\si{\per\cubic\centi\metre} (typical of microsecond-delay LIBS) to $5 \times 10^{17}$~\si{\per\cubic\centi\metre} (early-time, dense plasmas where Stark broadening dominates the line profiles).
% Source: element_id_benchmark_report.md §2.3
Each basis spectrum is computed with ionization stages I and II using exact partition functions evaluated directly from a database of 9{,}626~atomic energy levels from the NIST Atomic Spectra Database \citep{Kramida2024}, covering 83~elements across all relevant ionization stages.
This approach avoids reliance on polynomial partition function fits, which are available for only a small subset of elements, and follows the methodology of \citet{Barklem2016} who demonstrated that direct summation over experimental energy levels provides more accurate partition functions than polynomial approximations for the full periodic table.
Spectra are convolved with a Gaussian instrument profile (FWHM~$= 0.5$~nm, matched to RP~$\approx 600$--$1000$), and area-normalized to unit integral.
The resulting library is stored in HDF5 format with shape $(76 \times 300 \times 4096)$.
GPU-accelerated basis generation on the V100S completes in 3.1~minutes, compared to ${\sim}45$~minutes on CPU, enabling rapid iteration over basis library configurations during development.

\subsubsection{Identification Pathways}
\label{sec:id-pathways}

We evaluate five algorithmically distinct pathways, each with a systematic parameter sweep over its configuration space:

\paragraph{ALIAS peak-matching (baseline).}
The ALIAS algorithm \citep{Labutin2013} identifies elements by matching detected experimental peaks against theoretical line positions from an atomic database.
For each candidate element, ALIAS detects peaks via second-derivative enhancement, auto-calibrates the wavelength axis (global shift and effective RP estimation), estimates the plasma temperature from Boltzmann plot fitting, and scores each element by a confidence level (CL) combining match quality, Boltzmann consistency, and statistical significance against chance coincidence.
An element is detected if its CL exceeds the detection threshold.
We swept 18~configurations over detection threshold~$\in \{0.02, 0.03, 0.05\}$, intensity threshold factor~$\in \{3.0, 3.5\}$, and chance window scale~$\in \{0.3, 0.4\}$.

\paragraph{Full-spectrum NNLS decomposition.}
Inspired by spectral unmixing approaches in remote sensing \citep{Wang2014,Lawson1974}, this pathway decomposes the observed spectrum as a non-negative linear combination of single-element basis spectra:
\begin{equation}
  I_\mathrm{obs}(\lambda) \approx \sum_i c_i \, B_i(\lambda; T, n_e) + P(\lambda),
  \label{eq:nnls}
\end{equation}
where $B_i$ are the basis spectra at estimated $(T, n_e)$, $c_i \geq 0$ are element coefficients enforced by the NNLS constraint \citep{Lawson1974}, and $P(\lambda)$ is a polynomial continuum.
The physical basis for this decomposition is the approximate linearity of emission intensity with respect to elemental concentration under LTE conditions: the emissivity of each line scales linearly with the number density of the emitting species, which is proportional to the elemental concentration through the Saha--Boltzmann population distribution at fixed $T$ and $n_e$.
An element is identified as present if its coefficient signal-to-noise ratio (SNR) exceeds a threshold.
We swept 45~configurations over detection SNR~$\in \{1.0, 1.5, 2.0, 2.5, 3.0\}$, continuum degree~$\in \{2, 3, 4\}$, and fallback temperature~$\in \{6000, 8000, 10{,}000\}$~K.

\paragraph{Hybrid NNLS+ALIAS (two-stage).}
This novel approach combines the complementary strengths of NNLS (global spectral context, robust to blending) and ALIAS (individual line validation, pattern-matching precision) in a two-stage architecture inspired by the hierarchical spectral discrimination pipelines used in ChemCam/SuperCam \citep{Wiens2013,Clegg2017} and proposed by \citet{Eum2021}:
In the first stage, full-spectrum NNLS decomposition is performed with a lenient detection threshold (SNR~$\geq 1.0$--$2.0$), exploiting the global Boltzmann emissivity envelope to produce a candidate element set with high detection sensitivity.
In the second stage, ALIAS peak-matching is applied with the search space restricted to the NNLS candidates, validating that each candidate element has corresponding individual-line evidence in the observed spectrum.
In \emph{intersection} mode, an element must pass both stages; in \emph{union} mode, either stage suffices.
We swept 18~configurations over NNLS SNR~$\in \{1.0, 1.5, 2.0\}$, ALIAS detection threshold~$\in \{0.03, 0.05, 0.10\}$, and mode~$\in \{\text{intersect}, \text{union}\}$.

\paragraph{Voigt deconvolution + ALIAS.}
This pathway addresses the root cause of ALIAS failure at low RP---spectral overlap---by pre-processing the spectrum with multi-peak Voigt deconvolution \citep{AlJalali2016,Dai2026} before running ALIAS.
The procedure comprises baseline subtraction, peak detection, grouping of nearby peaks (within $2 \times$~FWHM), multi-peak Voigt fitting per group, reconstruction of a cleaned spectrum, and ALIAS identification on the cleaned spectrum.
We tested 2~configurations with detection threshold~$\in \{0.03, 0.05\}$.

\paragraph{Forward-model concentration thresholding.}
This pathway applies NNLS decomposition with a concentration-based detection criterion: after fitting, each element's coefficient is normalized to a fractional concentration ($c_i / \sum_j c_j$), and elements above a threshold are detected.
We swept 10~configurations over concentration threshold~$\in \{0.001, 0.005, 0.01, 0.02, 0.05\}$ and continuum degree~$\in \{2, 3\}$.

\paragraph{GPU-accelerated multi-method consensus identification optimization.}
In addition to the five primary pathways, we optimize a 101-parameter ensemble that fuses the outputs of all five identification algorithms via learned weights and per-element detection thresholds.
The 101 parameters span 8~groups: 5~algorithm selection weights (combined via softmax normalization), per-algorithm configuration tuning (ALIAS: 14~parameters, NNLS: 5, Hybrid: 13, Correlation: 17, Comb: 14), 22~per-element detection threshold multipliers, and 10~evidence combination parameters controlling signal-to-noise gating and confidence aggregation.
% Source: full experimental campaign, 101-param CMA-ES optimization
The optimization uses Separable CMA-ES \citep{Hansen2001,Ros2008} as implemented in the evosax library \citep{Lange2023}, a gradient-free evolutionary strategy well suited to non-differentiable objective functions.
The objective is the $F_{\beta=0.5}$ score \citep{VanRijsbergen1979} (precision-weighted, emphasizing false-positive suppression over recall), evaluated against precomputed cached outputs of all five algorithms on all 74~Aalto spectra.
Each candidate configuration is evaluated by applying exponential-simplex-weighted method selection, per-element identification confidence thresholds, and evidence combination logic to the cached method outputs, enabling a throughput of ${\sim}45{,}000$ configurations per second on a single NVIDIA V100S GPU with a population of 4{,}096 per generation.
A total of $2{,}048{,}000$ configurations were evaluated in 500~generations (58~seconds wall time), with convergence (gradient magnitude $< 0.003$) reached by generation~$\sim$300.
We interpret this optimization as systematic exploration of element-specific identification criteria, where each of the 101~parameters has spectroscopic meaning: method weights reflect the relative reliability of different spectral information channels (global emissivity fitting vs.\ individual line-position matching vs.\ multiplet spacing recognition), and per-element thresholds encode the spectroscopic difficulty of identifying each element at the given resolving power.
